---
title: "Reproducible_research_Assignment1"
output: html_document
keep_md: true 
---

# Downloading of Data

```{r echo=TRUE}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

if(!dir.exists("./dataActivity.zip")){
        download.file(fileUrl, "./dataActivity.zip", mode = "wb")
        unzip("./dataActivity.zip", exdir = )
}
```

# Loading and preprocessing the data

Show any code that is needed to:

1. Load the data (i.e. read.csv())

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r echo =TRUE}
# load data
activitydata <- read.csv("activity.csv", header = T)

# print out first 20 rows
head(activitydata,20)
```


# What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

1. Calculate the total number of steps taken per day

2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

3. Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE}
#What is mean total number of steps taken per day?
stepsSplit <- aggregate(steps ~ date, data = activitydata, sum, na.rm = T)

#Create histogram
hist(stepsSplit$steps)

# Calculate mean
m <- mean(stepsSplit$steps)

# Calulate Median
me <- median(stepsSplit$steps)

```

The mean is `r m`

The Median is `r me`

# What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}

# What is the average daily activity pattern?
stepbyinterval <- aggregate(steps ~ interval, activitydata, mean, na.rm=T)

# Create a time plot
plot(stepbyinterval$interval, stepbyinterval$steps,type = "l",
     xlab = "Average Number of Steps",
     ylab = "Interval")


# find row with max number of steps
maxStep <- which.max(stepbyinterval$steps)

# find the interval which consist of maximum number of steps
maxsteprow <- stepbyinterval[maxStep, ]

# Max step Interval details
maxsteprow

```

The Max Number of steps is `r maxsteprow$steps`

# Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


```{r echo=TRUE}
# Checking the number of NA values
NaSum <- sum(is.na(activitydata))
NaSum

data_imputed <- activitydata
for (i in 1:nrow(data_imputed)) {
        if (is.na(data_imputed$steps[i])) {
                interval_value <- data_imputed$interval[i]
                steps_value <- stepbyinterval[
                        stepbyinterval$interval == interval_value,]
                data_imputed$steps[i] <- steps_value$steps
        }
}

# Check if there is any NA Value
sum(is.na(data_imputed))

# imputed DF total number of step taken per interval
imputedTotalSteps <- aggregate(steps ~ date, data_imputed, sum)
head(imputedTotalSteps)

# create histogram
hist(imputedTotalSteps$steps, main = "Total number of step per day",
     xlab="Total number of steps in a day")

# Mean and Median of imputed data
mean(imputedTotalSteps$steps)

#median
median(imputedTotalSteps$steps)

#Mean and Median of orignal Data
mean(stepsSplit$steps)

#median
median(stepsSplit$steps)


```

# Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.


```{r echo=TRUE}
# Are there differences in activity patterns between weekdays and weekends?

data_imputed$dayType <- weekdays(as.Date(data_imputed$date))

data_imputed$dayType[data_imputed$dayType %in% c("Saturday", "Sunday")] <- "weekend"
data_imputed$dayType[data_imputed$dayType != "weekend"] <- "weekday"

data_imputed$dayType <- as.factor(data_imputed$dayType)

# calculate average steps by interval across all days
imputed_steps_by_interval <- aggregate(steps ~ interval + dayType, data_imputed, mean)

# make the panel plot for weekdays and weekends
library(lattice)

# create the panel plot
xyplot(steps ~ interval | dayType, imputed_steps_by_interval, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")
```

